---
title: "p8130_hw4_brm2150"
author: "Brooklynn McNeil"
date: "2024-11-17"
output: github_document
---

Brooklynn McNeil's homwork assignment 4 for p8130 Biostatistical Methods 1.

```{r setup, include=FALSE}
library(tidyverse)
library (BSDA)

knitr::opts_chunk$set(
  comment = '', fig.width = 8, fig.height = 6, out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 1

A new device for evaluating blood sugar levels is being tested.

a.    Below is the sign test for the blood sugar readings. The null hypothesis is that the median is equal to 120. The alternative hypothesis is one-sided that the blood sugar median is less than 120. Since the p-value is greater than 0.05 we fail to reject the null and conclude that the median is less than 120. 

```{r}
sugar_test = c(125,123,117,123,115,112,128,118,124,111,116,109,125,120,113,123,112,118,121,118,122,115,105,118,131)

SIGN.test(sugar_test, md = 120, alternative = "less", conf.level = 0.95) 
```
b.    Let's do the wilcoxon sign test with the same dataset.The p value from this test is also greater than 0.05, so we fail to reject the null again.

```{r}
wilcox.test(sugar_test, alternative = "less", mu = 120, conf.int = 0.95)
```

